SAT Competition 2020\tnoteref{title}

### 2020-12-27 paper/main.tex
planning and similarity

\todo{Similarity of Solvers: Rank Correlation}

@@

\subsubsection{CaDiCaL Hacks: Trail and AllUip}

This modified version of last years CaDiCaL was submitted with the two new methods \emph{Trail Saving}~\cite{Hickey:2020:TrailSaving} and \emph{Stable AllUIP}~\cite{Bacchus:SC2020}. 
It was submitted in three configurations, one with both methods activated and two other with each having only one of them active. 
The \emph{Stable AllUIP} variants seemed most successful, in most tracks they were the highest ranked variants and won the first price in the Planning Sub-Track. 
However, in the Unsat Sub-Track, the \emph{Trail Saving} variant performed best.

With \emph{Trail Saving}, the solver caches backtracked slices of the assignment trail in order to reuse them when the same decisions are repeated. 
The authors came up with sophisticated control mechanisms on when and how to reuse the cached trail slices and when to flush that cache.

With \emph{Stable AllUip}, the solver continues resolution beyond the 1st Unit Implication Point (1-UIP) during clause-learning~\cite{Zhang:2001:ClauseLearning} as long that does not increase the LBD value of the learnt clause and then keeps that clause only if it has a smaller size than then 1-UIP clause. 
The authors came up with an additional heuristic to limit the amount of attempted AllUip learning attempts, if the fraction of unsuccessful alluip learning attempts gets to high, which they check and adapt on every restart~\cite{Bacchus:SC2020}.

\subsubsection{Relaxed MapleLCMDistCBT-DL newTech}

The variant \emph{newTech} of the \emph{Relaxed} fork of the MapleLCMDistCBT-DL was especially successful on satisfiable instances. 
The solver won two awards, second price in Main Track and first price in its Sat Sub-Track.

The \emph{Relaxed} fork of MapleLCMDistCBT-DL already participated in SAT Race 2019~\cite{Xindi:SC2019}.
They integrated the local search solver CCAnr by extending the current partial assignment of the CDCL solver to a full assignment and for a fixed interval start the local search solver to try to find a nearby solution.

In the new versions of the Relaxed family of Maple solvers~\cite{Xindi:SC2020}, they probabilistically switch between several modes of phase selection, and some of these modes reuse the maximal partially satisfying assignments as determined by the local search solver.

For their successful \emph{newTech} variant~\cite{Xindi:SC2020} of their fork, they count the number of variable-occurrences in unsat clauses for each flip in the local search solver. 
Then they reuse this information to priorize variables in a modified branching heuristic.

\subsubsection{MapleLCMDistCBT F2TRC}

The MapleLCMDistCBT F2TRC forks were ranked third in the Unsat Sub-Track of the Main-Track.

The F2TRC forks aim at introducing determinism in winning solvers of previous competition, e.g., by replacing time-based intervals for LRB-VSIDS switches by a heuristic based on the conflict-count~\cite{Kochemazov:SC2020}.

Another modification aims at improving three-tier clause management based on the three tiers core, tier2 and local as it was introduced by Chanseok Oh~\cite{Oh:2015:satunsat}.
In order to limit growth of the core-tier, they introduce a limit on the number of core clauses, and whenever that is reached some of the low-ranked inactive core clauses are moved from core to tier2 and the limit is increased. 
In order to prevent tier2 from starving, they replaced the conflict limit for tier2 reduction by a size limit, then only half of the most recently inactive clauses is moved to local tier~\cite{Kochemazov:SC2020}.

@@

\begin{table}[ht]
\smaller
\centering
\begin{tabular}{|l|l|l|}
\hline
\bf Team & \bf Solver & \bf Awards\\
\hline
\multirow{2}{*}{\stack{Vallade, Le Frioux, Baarir, }{Sopena, Kordon}}~
 & P-MCOMSPS-STR-32 & \firsto \firsts \secondu \\
 & P-MCOMSPS-STR-64 & \firsto \\
\hline
\multirow{2}{*}{\stack{Biere, Fazekas, }{Fleury, Heisinger}}
 & Plingeling & \secondo \thirds \firstu \\
 & Treengeling & \\
\hline
\multirow{2}{*}{Nabeshima, Inoue}
 & ManyGlucose-32 & \thirdo \thirdu \\
 & ManyGlucose-64 & \thirdu \\
\hline
\multirow{2}{*}{Tchinda, Djamegni}
 & painlessmaplevone & \seconds \\
 & painlessmaplevtwo & \seconds \\
\hline
Li, Wu, Xu, Chen & syrupscavel & \\
\hline
Chen & abcdsatptwenty & \\
\hline
\end{tabular}
\caption{Teams and Solvers participating in the parallel track.}
\end{table}

>>>>

\subsection{Planning Sub-Track}
% - 86 out of 100 biggest (number of clauses) are planning
% - 109 / 200 in the planning track are not solved by any solver
    % - 186 / 600 in total
% - all 200 above 90\% binary clauses and half have above 99%
The instances of the planning sub-track are big compared to the rest of the
main track. Using the number of clauses as a metric; out of the $100$ biggest
instances
% of the main-track
$86$ belong to the planning track. The large size can mainly be attributed to
binary clauses. On average, more than $98$\% of the clauses are binary for
planning instances, where the average for the rest of the instances from the
main track is below $60$\%.

Despite the large size, only a few instances directly caused memory problems for
more than just a few solvers. In most cases the \SI{128}{\giga\byte} of RAM
available were sufficient.
% The saving grace might have been
% This can be attributed to
% the \SI{128}{\giga\byte} of RAM available to each solver.
However, the three biggest instances from the \emph{childsnack} family of
planning problems encoded with the \emph{hierarchical} planner \emph{Tree-REX}
caused over half of the solvers to abort with some error message related to
memory. Other solvers were still able to solve the same instances within $300$
seconds.

\subsection{Similarity of Solvers}
To investigate how similar the solvers of the main track are to each other, we
define a similarity metric based on the measured runtimes. For that we first
remove $186$ benchmarks from the collected data that have not been solved by any
solver. For the rest of the benchmarks we use the PAR-2 score,
% based on the \num{5000}-second timeout
i.\,e., for each solver we set a score of \num{10000} for unsolved instances.
This leaves us with $414$ floats
$S_{1},\dots,S_{n}$ for each solver $S$. The similarity for two solvers $A$
and $B$, normalized to the interval $[0,1]$ is defined
% as:
by Equation~\ref{eq-similarity}.

\begin{equation}
\label{eq-similarity}
	similarity(A, B) =\quad 1 - \frac{\sum{|A_{i} - B_{i}|}}{414 \cdot 10000}
\end{equation}
We calculated the \emph{similarity} between the $30$ solvers which achieved the
best PAR-2 score in the main track. The results are displayed in
Figure~\ref{fig-similarity-main} as a \emph{heat map}. Additionally the result
of hierarchically clustering the solvers based on their similarity is
illustrated as a \emph{dendrogram}. The height at which two solvers or clusters
are joined reflects how similar they are. For example
enabling the \emph{trail saving} in \emph{CaDiCaL-alluip} has no impact on the
runtime, resulting in a similarity above \num{0.999}. Thus the two solvers are joined very low in the dendrogram.
% A similar visualization was used in the SAT Challenge 2012 \todo{cite}. They used

One big cluster we can identify are the \emph{Mapel} decedents in the lower left
corner, all of which are modifications of the winners of the SAT Competition
2018 and the SAT Race 2019. \todo{I think... also race vs competition? Sorry I am new
  here NF} The solver \emph{mergesat} does not quite fit into the cluster and
has a comparatively high similarity to the \emph{CryptoMiniSat} (\emph{CMS})
family at the top right. With the exception of \emph{mergesat} and maybe
\emph{DurianSat} we observe a high similarity within the cluster.

The three \emph{Relaxed} solvers by \emph{Zhang} and \emph{Cai} are using the
same code base as a lot of the solvers in the Maple-Cluster but their overall
performance is better and closer to the CMS-Cluster.

The CMS versions differ in their implementation of stochastic local search (SLS)
but overall have a similar performance.

The last big cluster in the center of the heat map is comprised of the solvers
originally written by Biere. The three configurations of \emph{Kissat} form a
subcluster, even though the configurations are fairly distinct from each other.
% , specialized for satisfiable and unsatisfiable instances.
% One targets satisfiable instances, while the other is focused on unsatisfiable
% instances and a third alternating between them.

The other subcluster consists of
modifications to the 2019 version of \emph{\mbox{CaDiCaL}}. The new version
CaDiCaL-2020 has a worse overall performance than
% has an overall worse per
% is lacking behind
the other solvers in this cluster.

\begin{figure}[tp]
    \centering
    \includegraphics[width=\linewidth]{similarity/cross.pdf}
    \caption{Heat-map and dendrogram (top) based on the runtime similarity of the solvers participating in the main track.
      % Similarity is defined as
      % one minus the average taxicab distance normalized to the interval [0,1].
      % Spearman's rank correlation coefficient.
      Darker regions mean that the solvers are more similar.
      A more precise relation between color and similarity-value together with a histogram of the values that appear is given at the bottom.
    }
    \label{fig-similarity-main}
\end{figure}

% To estimate the similarity between the performance of the solvers of the Main
% Track we define a similarity metric $s$.
% identify each solver with a point in the 600-dimensional real vector
% space. Each component describes the PAR-2 runtime on one instance. We map the
% taxicab distance to the $[0, 1]$ interval and then to determine how similar two
% solvers are.

@@

\subsubsection{Trail Saving and Stable AllUip in CaDiCaL}

Based on \solver{CaDiCaL}, its variants \solver{Trail} and \solver{AllUip} present implementations of a new \emph{Trail Saving} approach~\cite{Hickey:2020:TrailSaving} and improved clause-learing based on the notion of \emph{Stable AllUIP}~\cite{Bacchus:SC2020}. 
Submitted were the three variants \solver{Trail}, \solver{AllUip} and \solver{AllUip+Trail}. 
The variants including \emph{Stable AllUIP} were the most successful and won the first price in the Planning Sub-Track.

The \emph{Stable AllUip} variants resolve additional clauses beyond the 1st Unit Implication Point (1-UIP) and keep them if they are of smaller size or LBD than the 1-UIP clause. 
By monitoring the frequency of clauses which successfully pass that filter, the solver dynamically limits the amount of attempts~\cite{Zhang:2001:ClauseLearning,Bacchus:SC2020}.
The \emph{Trail Saving} variant caches backtracked portions of the trail and uses them to restore decision levels during search if possible~\cite{Hickey:2020:TrailSaving}.

\subsubsection{Relaxed Maple newTech}

The \solver{Relaxed} fork of \solver{MapleLCMDistCBT-DL} was first presented to the community in SAT Race 2019~\cite{Xindi:SC2019}. 
In SAT Competition~2020, its new variant \solver{newTech} showed a good performance especially on satisfiable instances. 
The solver won two awards, the second price in the Main track and the first price in the Sat track.

The \solver{Relaxed} fork integrates short runs of the local search solver \solver{CCAnr} through periodic ex- and import of assignments~\cite{Xindi:SC2019}. 
The submitted versions of \solver{Relaxed Maple} use a probabilistic schedule for switching between $10$ phase selection modes. 
The award winning variant \solver{newTech} uses occurrence counts of variables in unsatisfied clauses during SLS runs to recalculate variable priorities for their modified branching heuristic~\cite{Xindi:SC2020}.

\subsubsection{Maple F2TRC}

The \solver{F2TRC} fork of \solver{MapleLCMDistCBT} won the third price in the Unsat track. 
\solver{F2TRC} comes with a deterministic reimplementations of former winning strategies, e.g., by replacing time intervals by conflict intervals~\cite{Kochemazov:SC2020}.

\solver{F2TRC} introduces an improved management of learned clauses in the tree tiers \emph{core}, \emph{tier2} and \emph{local} which are inherited from \solver{CoMinisatPS}~\cite{Oh:2015:satunsat}.
A dynamic size limit for the \emph{core} tier, triggers the reassignment of inactive clauses from \emph{core} to \emph{tier2}. 
To conter-act an observered starving of tier2, the conflict-based heuristic that controls demotion of clauses from \emph{tier2} to \emph{local} was replaced by a size based heuristic~\cite{Kochemazov:SC2020}.

@@

\begin{table}[b]
\smaller
\centering
\begin{tabular}{|l|l|l|}
\hline
\bf Team & \bf Solver & \bf Awards\\
\hline
\multirow{2}{*}{\stack{Vallade, Le Frioux, Baarir, }{Sopena, Kordon}}~
 & P-MCOMSPS-STR-32 & \firsto \firsts \secondu \\
 & P-MCOMSPS-STR-64 & \firsto \\
\hline
\multirow{2}{*}{\stack{Biere, Fazekas, }{Fleury, Heisinger}}
 & Plingeling & \secondo \thirds \firstu \\
 & Treengeling & \\
\hline
\multirow{2}{*}{Nabeshima, Inoue}
 & ManyGlucose-32 & \thirdo \thirdu \\
 & ManyGlucose-64 & \thirdu \\
\hline
\multirow{2}{*}{Tchinda, Djamegni}
 & painlessmaplevone & \seconds \\
 & painlessmaplevtwo & \seconds \\
\hline
Li, Wu, Xu, Chen & syrupscavel & \\
\hline
Chen & abcdsatptwenty & \\
\hline
\end{tabular}
\caption{Teams and Solvers participating in the parallel track.}
\end{table}
